---
title: "Native_vs_Tx_Fibrosis"
output: html_document
date: "2023-11-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will need our TOP package from Bioc.
```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("TOP")
```

Load in the libraries we need for this.
```{r}
library(TOP) # Version 1.2.0
library(dplyr)
library(tibble)
```

First we 
```{r}
setwd("/dskh/nobackup/harryr/Pan-Organ Allograft Dysfunction/Fibrosis/Processed_Datasets")
files <- list.files(pattern = "_Tx")
Fibrosis_Counts <- lapply(files, function(x){
  y <- read.csv(x) 
  y %>%
    dplyr::select(-X) %>%
    tibble::remove_rownames() %>%
    filter(!duplicated(Gene.Symbol) & !is.na(Gene.Symbol)) %>%
    tibble::column_to_rownames("Gene.Symbol") %>%
    dplyr::select(t)
})

# We only want to keep datasets that have a meaningful number of features.
keep <- Fibrosis_Counts %>% sapply(nrow) > 10000 
Fibrosis_Counts <- Fibrosis_Counts[keep]
int <- Reduce(intersect, lapply(Fibrosis_Counts, rownames))
test_list <-  lapply(Fibrosis_Counts, function(x)x[int,, drop = FALSE])

# concatinate the gene level summaries across datasets.
test_df <- do.call(cbind, test_list)
Z.Scores.All <- apply(test_df, 2, function(x) {
        stats::qnorm(rank(x)/(nrow(test_df) + 1))
})

# Perform pvalue combination using the directPA package.
data(Pathways, package = "directPA")
gene.pvalues <- apply(Z.Scores.All, 1, function(x) {
  directPA::geneStats(x)
})
gene.zscores <- stats::qnorm(gene.pvalues, lower.tail = FALSE)
pvalue2sided <- 2 * stats::pnorm(-abs(gene.zscores))

# Let's pull out the top 50 genes.
sig.genes <- names(sort(pvalue2sided))[1:50]
```

Let's compare the gene expression between native and transplant organ fibrosis.
```{r}
setwd("/dskh/nobackup/harryr/Pan-Organ Allograft Dysfunction/Fibrosis/Processed_Datasets")

# Read in the data
names <-list.files()
list_tStats <- lapply(list.files(), read.csv)
list_tStats <- lapply(list_tStats, function(x){
  x %>%
    dplyr::select(Gene.Symbol, t)
})

# Pull everything into one clean dataframe
list_tStats <- Reduce(function(x, y) merge(x, y, by = "Gene.Symbol", all = TRUE), list_tStats)
list_tStats <- list_tStats[which(!duplicated(list_tStats$Gene.Symbol) & !is.na(list_tStats$Gene.Symbol)), ]
list_tStats <- list_tStats[-c(1:3),] # These genes make no sense
rownames(list_tStats) <- NULL

# Remove any NA's
library(tibble)
list_tStats <- list_tStats %>%
  column_to_rownames(var = "Gene.Symbol") %>%
  na.omit()

# Annotating dataframe
library(TOP)
GSE <- TOP:::str_split_n(names, "_", 2)
OrganTx <- TOP:::str_split_n(names, "_", 5)
TxType <- sub(".csv", "", TOP:::str_split_n(names, "_", 6))

anno_df <- data.frame(OrganTx, TxType)

# Seperating Native and Tx
list_tStats_Tx <- list_tStats[,which(TxType == "Tx")]
list_tStats_Native <- list_tStats[,which(TxType != "Tx")]

# Run directPA
library(directPA)
library(PhosR)
data(Pathways)

# Combined pvalues using directPA
Z.Scores.All <- apply(list_tStats_Tx, 2, function(x){qnorm(rank(x)/(nrow(list_tStats_Tx)+1))})
gene.pvalues <- apply(Z.Scores.All, 1, function(x){geneStats(x)})
gene.zscores_Tx <- qnorm(gene.pvalues, lower.tail = FALSE) %>%
  data.frame()

# Combined pvalues using directPA
Z.Scores.All <- apply(list_tStats_Native, 2, function(x){qnorm(rank(x)/(nrow(list_tStats_Native)+1))})
gene.pvalues <- apply(Z.Scores.All, 1, function(x){geneStats(x)})
gene.zscores_Native <- qnorm(gene.pvalues, lower.tail = FALSE) %>%
  data.frame()

# Merge the two results into one dataframe.
all(rownames(gene.zscores_Tx) == rownames(gene.zscores_Native))
df_zscore <- cbind(gene.zscores_Tx, gene.zscores_Native)
colnames(df_zscore) <- c("Tx", "Native")

# Plot Results
library(plotly)
data(Pathways)
dPA <- directPA(df_zscore, direction = pi/4, annotation = Pathways.reactome, visualize = TRUE, 
                        ylab = "Native Organ Fibrosis", 
                        xlab = "Transplant Organ\nFibrosis")
```

Next, load in the AUSCAD data for validation.
```{r, warning=FALSE, message=FALSE}
DGE <- readRDS("/dskh/nobackup/harryr/POAD_Validation/AUSCAD_3m_Bx_Fibrosis.Rds")
lcpm <- DGE$counts # Pull out counts.
```

We will now load in the fibrosis counts. Eventually returning a list containing a matrices of counts, and a list containing the different outcomes for each dataset.
```{r}
setwd("/dskh/nobackup/harryr/Pan-Organ Allograft Dysfunction/Fibrosis/Normalised_Counts/")
files <- list.files()
Biopsy_Counts <- lapply(files, function(x){
  y <- read.csv(x) 
})

Biopsy_Counts <- lapply(Biopsy_Counts, function(x){
  x %>%
    dplyr::select(-X) %>%
    tibble::remove_rownames() %>%
    filter(!duplicated(Gene_Symbol) & !is.na(Gene_Symbol)) %>%
    tibble::column_to_rownames("Gene_Symbol")
})

int <- Reduce(intersect, lapply(Biopsy_Counts, rownames))
test_list <-  lapply(Biopsy_Counts, function(x)x[int,, drop = FALSE])

y_list <- list()
for(i in seq_along(test_list)){
  names <- colnames(test_list[[i]])
  y_list[[i]] <- factor(ifelse(grepl("Control", names), "Control", "Fibrosis"))
}
test_list <- lapply(test_list, t)
```

We now want to build models based on naive or transplant fibrosis. To do so, we need to pull out dataset specific naming from the initial file naming.
```{r}
setwd("/dskh/nobackup/harryr/Pan-Organ Allograft Dysfunction/Fibrosis/Processed_Datasets")
names <- list.files()

# Annotating dataframe
library(TOP)
GSE <- TOP:::str_split_n(names, "_", 2)
OrganTx <- TOP:::str_split_n(names, "_", 5)
TxType <- sub(".csv", "", TOP:::str_split_n(names, "_", 6))

# merge the organ type and transplant (Tx or Naive) into one df.
anno_df <- data.frame(OrganTx, TxType)
```

Next, we will use TOP to build a transplant specific fibrosis model.
```{r}
library(TOP)

# Pull out all the datasets that are from transplant (Tx) fibrosis.
tx_counts <- test_list[TxType == "Tx"]
tx_outcome <- y_list[TxType == "Tx"]

tx_features <- TOP::filterFeatures(tx_counts, tx_outcome, 
                                   contrast = "Fibrosis - Control", nFeatures = 100)
tx_features <- intersect(tx_features, rownames(lcpm))
tx_counts = lapply(tx_counts, "[", , tx_features)

tx_model <- TOP::TOP_model(tx_counts[-9], tx_outcome[-9]) # the 9th element of the list had too much missing data to be included.

# Make predictions on the AUSCAD cohort. 
filt_lcpm <- lcpm[tx_features,]
tx_fib_predictions <- TOP::predict_TOP(tx_model$models, as.matrix(t(filt_lcpm)))
DGE$samples$tx_fib_predictions <- as.numeric(tx_fib_predictions)
```


Next, we will use TOP to build a native-organ fibrosis model.
```{r}
library(TOP)

# Pull out all the datasets that are from naive fibrosis.
native_counts <- test_list[TxType == "Native"]
native_outcome <- y_list[TxType == "Native"]

# the 1st element of the list had too much missing data to be included.
tx_features <- TOP::filterFeatures(native_counts[-1], native_outcome[-1], 
                                   contrast = "Fibrosis - Control", nFeatures = 100)
tx_features <- intersect(tx_features, rownames(lcpm))
tx_counts = lapply(native_counts, "[", , tx_features)

# the 1st element of the list had too much missing data to be included.
tx_model <- TOP::TOP_model(tx_counts[-1], native_outcome[-1])

# Make predictions on the fibrosis dataset from AUSCAD.
filt_lcpm <- lcpm[tx_features,]
native_fib_predictions <- TOP::predict_TOP(tx_model$models, as.matrix(t(filt_lcpm)))
DGE$samples$native_fib_predictions <- as.numeric(native_fib_predictions)
```
We can create a custom plot that has the colours that we want.
```{r}
myROC_Plot <- function (roc_list) 
{
  library(pROC)
    data.labels <- TOP:::extractAUC(roc_list)
    pROC::ggroc(roc_list, size = 1.5) + theme_bw() +
  scale_color_manual(name = "Model", values = c("#FBA730", "#0A87C6", "#FF9DA7"),
                     labels = data.labels$label_long) +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color = "grey50", linetype = "dashed") +
  theme(legend.title = element_text(size = 14)) +
  theme(legend.text = element_text(size = 12)) +
  ggtitle("")
}
```

Plot the transplant and naive models. 
```{r}
myROC_Plot(list(`Tx Model` = roc(DGE$samples$fibrosis_bin  ~ DGE$samples$tx_fib_predictions),
                  `Native Model` = roc(DGE$samples$fibrosis_bin  ~ DGE$samples$native_fib_predictions)))
```

```{r}
sessionInfo()
```



